Adaptive Safe Strong Screening for Efficient Lasso-Type Problems Optimization over the Solution Path

### 2021-01-28 batch-hssr.tex
Updates from Overleaf
Edits to 3.1

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[ruled,vlined, linesnumbered]{algorithm2e}
\usepackage[english]{babel}
\usepackage[nottoc]{tocbibind}
\usepackage{color}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{float}
\newtheorem{theorem}{Theorem}[section]
\bibliographystyle{ims}
% Custom commands / shortcuts
\providecommand{\sign}{\textrm{sign}}
\providecommand{\pb}[1]{\textcolor{red}{#1}}
\providecommand{\lam}{\lambda}

@@

\begin{abstract}
Lasso type models are popular in statistics and machine learning, especially for variable selection in high-dimensional data. Typical practice is to tune the size of lasso penalty along a path of values. Due to modern data collection techniques, researchers need to analyse ultra high-dimension data with millions of features, which brings a great need for an efficient lasso optimization algorithm. Feature screening techniques have proven to be powerful for this need, because they can discard features and lead to a model with much less features. In this paper, we develop an adaptive hybrid screening framework where screening is carried out adpatively along the path of tuning parameter values. It reuses previous solutions in the path and avoids unhelpful heavy computations. It is a flexible framework that can be easily extended to different types of lasso models. The Ada-SSR-EDPP for the standard lasso model and the Ada-SSR-Slores for the sparse logistic model are shown as examples. We perform simulation study and real data study in a wide range of scenarios and the adaptive hybrid methods outperform other state-of-the-art method significantly and uniformly, with the greatest speedup in the most challenging scenarios.
\end{abstract}

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[ruled,vlined, linesnumbered]{algorithm2e}
\usepackage[english]{babel}
\usepackage[nottoc]{tocbibind}
\usepackage{xcolor}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{float}
\usepackage{hyperref}
\hypersetup{colorlinks=true, citecolor=black, linkcolor=., urlcolor=cyan}
\newtheorem{theorem}{Theorem}[section]
\bibliographystyle{ims}
% Custom commands / shortcuts
\providecommand{\sign}{\textrm{sign}}
\providecommand{\pb}[1]{\textcolor{red}{#1}}
\providecommand{\lam}{\lambda}

@@

\begin{abstract}
Lasso type models are popular in statistics and machine learning, especially for variable selection in high-dimensional data. Typical practice is to tune the size of lasso penalty along a path of values. Due to modern data collection techniques, researchers need to analyse ultra high-dimension data with millions of features, which brings a great need for an efficient lasso optimization algorithm. Feature screening techniques have proven to be powerful for this need, because they can discard features and lead to a model with much less features. In this paper, we develop an adaptive hybrid screening framework where screening is carried out adpatively along the path of tuning parameter values. It reuses previous solutions in the path and reduces unhelpful heavy computations. It is a flexible framework that can be easily extended to different types of lasso models. Two applications for the standard lasso model and the sparse logistic model are shown as examples. We perform simulation study and real data study in a wide range of scenarios and the adaptive hybrid methods outperform other state-of-the-art method significantly and uniformly, with the greatest speedup in the most challenging scenarios.
\end{abstract}

### 2021-01-28 main.tex

\begin{figure}[h]
    \centering
    \includegraphics[scale = 0.59]{plots/512.pdf}    \caption{Comparing speed of screening methods for sparse logistic regression model under different settings. Top row: computation time in second. Bottom row: relative computation time compared to SSR. First column: varying number of features $p$. Second column: varying sample size $n$. Third column: varying signal strength $\beta_{\max}$. Fourth column: varying sparsity $s$.}
    \label{fig:5.1.2a}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[scale = 0.59]{plots/512b.pdf}    \caption{Comparing speed of screening methods for sparse logistic regression model under different settings. Top row: computation time in second. Bottom row: relative computation time compared to SSR. First column: Varying number of grid points $L$. Second column: Varying range of grid $\lambda_{\min}/\lambda_{\max}$. Third column: Varying number of grid points $L$ with fixed spacing.}
    \label{fig:5.1.2b}
\end{figure}

Figure 4 and 5 show the computation time (top row) and relative computation time compared to SSR (bottom row) of solving sparse logistic regression problem with different screening methods. The overall pattern is the same as for sparse linear regression, with adaptive screening uniformly faster than all other methods. The speedup is smaller in magnitude than for linear regression because the Slores rule is less powerful than EDPP. Nevertheless, the improvement in speed is still dramatic in many cases. For example, when $L=1,000$, AHR was 4 times faster than any other method.

\subsection{Real data analysis}
\label{sec:real-data}

@@

For GENE and GWAS, we replicate the experiment 10 times to take the average. For data sets MNIST and NYT, because the response is randomly selected, we replicate the experiment 50 times to get an accurate estimate. Table 1 compares the average computation time for different screening methods in these data sets.

\begin{table}[H]
\centering
\begin{tabular}{l|l|l|l|l}
\hline
 & GENE & MNIST & GWAS & NYT \\
Screening method & n=536 & n=784 & n=313 & n=5,500 \\
 & p=17,322 & p=60,000 & p=660,496 & p=55,000 \\ \hline
AC & 2.14 (0.03) & 6.82 (0.07) & 46.67 (0.13) & 59.20 (1.89) \\
SSR & 1.34 (0.01) & 5.20 (0.01) & 25.34 (0.08) & 35.40 (0.44) \\
SEDPP & 1.38 (0.01) & 8.40 (0.45) & 25.48 (0.05) & 53.69 (3.81) \\
HSSR & 1.21 (0.01) & 3.44 (0.07) & 23.85 (0.06) & 31.05 (0.80) \\
AHR & \textbf{0.71 (0.01)} & \textbf{1.15 (0.02)} & \textbf{8.80 (0.03)} & \textbf{12.71 (1.05)} \\\hline
\end{tabular}
\caption{Average computing time in second (standard error) for solving the lasso model}
\end{table}

>>>>

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{plots/512.pdf}    \caption{Comparing speed of screening methods for sparse logistic regression model under different settings. Top row: computation time in second. Bottom row: relative computation time compared to SSR. First column: varying number of features $p$. Second column: varying sample size $n$. Third column: varying signal strength $\beta_{\max}$. Fourth column: varying sparsity $s$.}
    \label{fig:5.1.2a}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.82\textwidth]{plots/512b.pdf}    \caption{Comparing speed of screening methods for sparse logistic regression model under different settings. Top row: computation time in second. Bottom row: relative computation time compared to SSR. First column: Varying number of grid points $L$. Second column: Varying range of grid $\lambda_{\min}/\lambda_{\max}$. Third column: Varying number of grid points $L$ with fixed spacing.}
    \label{fig:5.1.2b}
\end{figure}

Figure~\ref{fig:5.1.2a} and~\ref{fig:5.1.2b} show the computation time (top row) and relative computation time compared to SSR (bottom row) of solving sparse logistic regression problem with different screening methods. The overall pattern is the same as for sparse linear regression, with adaptive screening uniformly faster than all other methods. The speedup is smaller in magnitude than for linear regression because the Slores rule is less powerful than EDPP. Nevertheless, the improvement in speed is still dramatic in many cases. For example, when $L=1,000$, AHR was 4 times faster than any other method.

\subsection{Real Data Analysis}
\label{sec:real-data}

@@

For GENE and GWAS, we replicate the experiment 10 times to take the average. For data sets MNIST and NYT, because the response is randomly selected, we replicate the experiment 50 times to get an accurate estimate. Table~\ref{Tab:real} compares the average computation time for different screening methods in these data sets.

\begin{table}[H]
\centering
\begin{tabular}{l|l|l|l|l}
\hline
 & GENE & MNIST & GWAS & NYT \\
Screening method & n=536 & n=784 & n=313 & n=5,500 \\
 & p=17,322 & p=60,000 & p=660,496 & p=55,000 \\ \hline
AC & 2.14 (0.03) & 6.82 (0.07) & 46.67 (0.13) & 59.20 (1.89) \\
SSR & 1.34 (0.01) & 5.20 (0.01) & 25.34 (0.08) & 35.40 (0.44) \\
SEDPP & 1.38 (0.01) & 8.40 (0.45) & 25.48 (0.05) & 53.69 (3.81) \\
HSSR & 1.21 (0.01) & 3.44 (0.07) & 23.85 (0.06) & 31.05 (0.80) \\
AHR & \textbf{0.71 (0.01)} & \textbf{1.15 (0.02)} & \textbf{8.80 (0.03)} & \textbf{12.71 (1.05)} \\\hline
\end{tabular}
\caption{Average computing time in second (standard error) for solving the lasso model}
\label{Tab:real}
\end{table}
